---
title: "ADSC - Case 2"
author: "Mint Thai - T00762325


Neha Malhan T00757554


Thiwanka Abeyweera T00754871"
date: "2024-09-23"
output: pdf_document
---

## Introduction (Mint)

An amateur investor seeks professional guidance on three proposed investment strategies for a \$100 investment. Each strategy presents a variety of potential outcomes, ranging from a loss of \$100 to a gain of up to \$5000, accompanied by their corresponding probabilities. This report will assess each strategy by calculating the expected value to determine potential profitability and analyze the associated risks through variance. The analyses and findings will be detailed in the methodology, results, and discussion sections, concluding with a recommendation for the investor's \$100 investment.

## Methodology (Thiwanka)

expected value

variance

this is a discrete random variable case

## Results (shared)

```{r, echo=FALSE}

# Strategy1
strategy1 <- data.frame(
  Outcome = c(-100, -50, 0, 10, 100, 200, 300, 400, 500),
  Probability = c(0.25, 0.3, 0.2, 0.1, 0.07, 0.04, 0.02, 0.01, 0.01)
)

# Strategy2
strategy2 <- data.frame(
  Outcome = c(-100, -75, -50, -25, 100, 200, 300, 400, 500, 1000),
  Probability = c(0.01, 0.05, 0.1, 0.25, 0.2, 0.2, 0.1, 0.05, 0.03, 0.01)
)

# Strategy3
strategy3 <- data.frame(
  Outcome = c(-100, -50, 0, 100, 1000, 2000, 3000, 4000, 5000),
  Probability = c(0.4, 0.3, 0.183, 0.1, 0.007, 0.004, 0.003, 0.002, 0.001)
)

# Calculate expected values for each strategy
ev_strategy1 <- sum(strategy1$Outcome * strategy1$Probability)
ev_strategy2 <- sum(strategy2$Outcome * strategy2$Probability)
ev_strategy3 <- sum(strategy3$Outcome * strategy3$Probability)


# Calculate variance for each strategy
variance_strategy1 <- sum((strategy1$Outcome - ev_strategy1)^2 * strategy1$Probability)
variance_strategy2 <- sum((strategy2$Outcome - ev_strategy2)^2 * strategy2$Probability)
variance_strategy3 <- sum((strategy3$Outcome - ev_strategy3)^2 * strategy3$Probability)

```

# Neha - strategy 1

# Thiwanka - strategy 2

# Mint - strategy 3

Figure X visualizes the outcomes and associated probabilities of Strategy 3 in a bar chart. Although this strategy offers a higher potential return of up to \$5000, it also comes with significant risks, as there is a 70% chance of losing money. The probability of breaking even is 20%, and the chances of making a profit are less than 12%.

```{r, echo = FALSE, fig.align='center',fig.cap="Strategy 3 Outcomes and Probabilities", out.width='50%'}

barplot(
  height = c(strategy3$Probability*100),
  names.arg = strategy3$Outcome,
  col = "lightblue",
  ylab = "Probability (%)",
  xlab = "Outcome ($)"
)

```

Based on the assessment and calculations, the expected value and variance for Strategy 3 are as indicated in Table X below:

```{r, echo = FALSE, fig.align='center',fig.cap="Strategy 3 Assessment Result", out.width='50%'}

strategy3_summary <- t(data.frame(
  ev_strategy3, variance_strategy3))

knitr::kable(strategy3_summary, digits = 0, caption = "Strategy 3 Assessment Result")

```

The **expected value** for Strategy 3 is **-8**. This indicates that the investor can expect an average loss of \$8 per investment. The negative expected value implies a greater likelihood of incurring losses rather than achieving gains and suggests that this strategy may not be favourable for the investor seeking profit. Additionally, the highest variance of 112686 suggests that the outcomes can vary widely, leading to extreme gains or losses. This highlights a significant risk and uncertainty associated with this investment strategy.

Given the negative expected value and high variance, Strategy 3 is a risky investment option for investors focused on financial stability and positive returns. This strategy may be more suitable for investors with a high-risk tolerance who are willing to accept the possibility of losses in pursuit of higher returns.

## Conclusions / Discuss (shared)

```{r echo=FALSE}

# Display Expected Value and Variance for all 3 strategies in a table

results_all <- data.frame(
  Strategy = c("Strategy 1", "Strategy 2", "Strategy 3"),
  Expected_Value = c(ev_strategy1, ev_strategy2, ev_strategy3),
  Variance = round(c(variance_strategy1, variance_strategy2, variance_strategy3)))

results_all_flip <- t(results_all[, -1])

knitr::kable(results_all_flip, col.names = c("Strategy 1", "Strategy 2", "Strategy 3"), digits = 0, caption = "Strategy Assessment Summary")

```

conclusion / discussion -- quickly sum everything up, write a brief comparison about the 3 strategies

## Recommendations (Neha)

## Appendices

```{r}

# Strategy1
strategy1 <- data.frame(
  Outcome = c(-100, -50, 0, 10, 100, 200, 300, 400, 500),
  Probability = c(0.25, 0.3, 0.2, 0.1, 0.07, 0.04, 0.02, 0.01, 0.01)
)

# Strategy2
strategy2 <- data.frame(
  Outcome = c(-100, -75, -50, -25, 100, 200, 300, 400, 500, 1000),
  Probability = c(0.01, 0.05, 0.1, 0.25, 0.2, 0.2, 0.1, 0.05, 0.03, 0.01)
)

# Strategy3
strategy3 <- data.frame(
  Outcome = c(-100, -50, 0, 100, 1000, 2000, 3000, 4000, 5000),
  Probability = c(0.4, 0.3, 0.183, 0.1, 0.007, 0.004, 0.003, 0.002, 0.001)
)

# Calculate expected values for each strategy
ev_strategy1 <- sum(strategy1$Outcome * strategy1$Probability)
ev_strategy2 <- sum(strategy2$Outcome * strategy2$Probability)
ev_strategy3 <- sum(strategy3$Outcome * strategy3$Probability)


# Calculate variance for each strategy
variance_strategy1 <- sum((strategy1$Outcome - ev_strategy1)^2 * strategy1$Probability)
variance_strategy2 <- sum((strategy2$Outcome - ev_strategy2)^2 * strategy2$Probability)
variance_strategy3 <- sum((strategy3$Outcome - ev_strategy3)^2 * strategy3$Probability)

```

```{r}

# Display Expected Value and Variance for all 3 strategies in a table

results_all <- data.frame(
  Strategy = c("Strategy 1", "Strategy 2", "Strategy 3"),
  Expected_Value = c(ev_strategy1, ev_strategy2, ev_strategy3),
  Variance = round(c(variance_strategy1, variance_strategy2, variance_strategy3)))

results_all_flip <- t(results_all[, -1])

knitr::kable(results_all_flip, col.names = c("Strategy 1", "Strategy 2", "Strategy 3"), digits = 0, caption = "Strategy Assessment Summary")

```

team work division
